---
title: "BDA - Assignment 7"
author: "Anonymous"
date: "29/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

```{r message=FALSE, warning=FALSE}
# Loading packages
library(tidyverse) 
library(aaltobda) # contains data "drowning" & "factory"
library(rstan)
# stan settings:
source('stan_utility.R') # diagnosis of rhats
options(mc.cores = parallel::detectCores()) #when using locally computer
rstan_options(auto_write = TRUE) # autosave compiled Stan program
# bay settings:
library(loo) #estimate predictive error of MCMC item-level log likelihood output
library(gridExtra)
library(bayesplot) #plots of posterior draws
library(shinystan) # model paramteres & MCMC simulations
#theme_set(bayesplot::theme_default(base_family = "sans"))
bayesplot_theme_set() #default
SEED <- 48927 # set random seed for reproducability
```

## Q1 - Linear model: drowning data with Stan

The drowning data contains the number of people drowned in Finland 1980-2016, which we want to investiagte using a linear model with Gaussian noise.

### Q1.1 - fixing errors
A Stan code is given with errors. We want to find and fix these.
The errors are marked in the following:

```{r eval=FALSE}
data {  
    int<lower=0> N; // number of data points 
    vector[N] x; // observation year 
    vector[N] y; // observation number of drowned 
    real xpred; // prediction year 
} 
parameters { 
  real alpha; 
  real beta; 
  real <upper=0> sigma; #error: variance can't be NEG
} 
transformed parameters { 
  vector[N] mu; 
  mu = alpha + beta*x; 
} 
model { 
  y ~ normal(mu, sigma); 
} 
generated quantities { 
  real ypred; 
  ypred = normal_rng(mu, sigma); #error: mu is a vector parametre
} 

```

For sigma error, we need to replace upper with lower.

As for "mu is a vector parameter": ypred has been defined as a real value, which will cause an error since mu is a vector parametre. 

After correcting the errors we get the following stan model:
```{r} 
writeLines(readLines("ex7_1.stan"))
```


### Q1.2 - Suitable values for $\tau$

No prior for the parametres is defined in "ex7_1.stan", which corresponds to using a uniform prior.

```{r}
writeLines(readLines("ex7_2.stan"))
```

```{r}
data("drowning")

d_lin <- list(N = nrow(drowning),
              x = drowning$year,
              y = drowning$drownings ,
              xpred = 2019)

fit_lin <- stan(file="ex7_2.stan", data = d_lin, seed = SEED,control = list(max_treedepth =15))
```

We would like to apply a weakly informative prior $\beta \sim \mathcal{N}(0,\tau^2)$ for the slope paramtere $\beta$.
Given is:
$Pr(-69<\beta<69) = 0.99$. 
Due to symmetry of the normal distribution, we get $p(\beta\le-69) = 0.005$, so tau can be calculated by:

```{r}
# tau calcuation
tau <- -69 / qnorm(0.005) 
tau
```

So we find tau to 26.8

### Q1.3 - Implementation of the prior

We add the calculated tau: 
```{r}
d_lin_prior <- c(list(
    tau = 26.8),
  d_lin)
```

Implementing in a new stan model:
```{r}
writeLines(readLines("ex7_3.stan"))
```

Fitting the model to the data
```{r}
fit_lin <- stan(file="ex7_3.stan", data = d_lin_prior, seed=SEED, control = list(max_treedepth =15))
```


```{r include=FALSE}
monitor(fit_lin, probs = c(0.1, 0.5, 0.9))

print(fit_lin)
check_hmc_diagnostics(fit_lin)

samples_lin <- rstan::extract(fit_lin, permuted = T)
 mean(samples_lin$beta>0) # probability that beta > 0
 
mu <- apply(samples_lin$mu, 2, quantile, c(0.05, 0.5, 0.95)) %>%
  t() %>% data.frame(x = d_lin$x, .)  %>% gather(pct, y, -x)
```

Visualizing to compare with plot form ex7:
```{r}
p <- ggplot() +
  geom_point(aes(x, y), data = data.frame(d_lin), size = 1) +
  geom_line(aes(x, y, linetype = pct), data = mu, color = 'red') +
  scale_linetype_manual(values = c(2,1,2)) +
  labs(y = 'number of drowned', x= "Year") +
  guides(linetype = F)

pars <- intersect(names(samples_lin), c('beta','ypred')) 
draws <- as.data.frame(fit_lin)
phist <- mcmc_hist(draws, pars = pars)
grid.arrange(p, phist, nrow = 2)
```

Comparing with the plot given in the exercise: Looks satisfying

## Q2 - Hierarchical model: factory data with Stan

## Seperated Gaussian Model

Implementing the seperated model:
```{r}
writeLines(readLines("ex7_4.stan"))
```

The data related to this model is : 
```{r}
data("factory")
data_separate <-list(N = 6*nrow(factory),
                     K = 6,
                     x = rep(1:6, nrow(factory)),
                     y = c(t(factory)))

#fitting in stan:
fit_sep <- stan(file="ex7_4.stan", data = data_separate, seed = SEED)
```


### The posterior distribution of the mean of the sixth machine:

```{r}
draws_separate <- as.data.frame(fit_sep)
mcmc_hist(draws_separate, pars = c("mu[6]", "ypred"))
```

### The posterior distribution of the mean of the quality measurements of the seventh machine:


### Pooled model

For the pooled model we assume that $\mu$ and $\sigma$ are the same for all machines. The Stan implementation for pooled model is as:
```{r}
writeLines(readLines("ex7_5.stan"))
```

The data related to this model is : 
```{r}
data_pooled <- list(N = 6*nrow(factory),
                    y = c(t(factory)))
```

We fit the pooled model in stan as follows: 
```{r}  
fit_pooled <- stan(file = "ex7_5.stan", data = data_pooled, seed = SEED)
```


## i) The posterior of the mean of the sixth machine:
$$ p(\mu_6|\sigma, y) \propto \mathcal{N}(\mu, \sigma^2) $$

## ii) The predictive distribution for another quality measurement from the sixth machine: 

$$ p(\hat{y_6}|\mu,\sigma) \propto  \mathcal{N}(\mu, \sigma^2) $$

## iii) The posterior distribution of the mean of the quality measurements of the seventh machine:
assuming the machines are identical, gives us: 
$$ p(\mu_7|\mu, \sigma) \propto \mathcal{N}(\mu, \sigma^2)$$
```{r}
draws_pooled <- as.data.frame(fit_pooled)
mcmc_hist(draws_pooled, pars = c("mu", "ypred", "mu_7"))
```



## Hierarchical model 

The Stan implementation of the model:

```{r}
writeLines(readLines("ex7_6.stan"))
```

We fit the separate model in stan: 
```{r}  
fit_hierarchical <- stan(file="ex7_6.stan", data = data_separate, seed = SEED)
```

## The predictive distribution for another quality measurement from the sixth machine: 


```{r} 
draws_hierarchical <- as.data.frame(fit_hierarchical)
mcmc_hist(draws_hierarchical, c("mu[6]", "ypred", "mu_7"))
```

**References:**

Code examples are found at Github:  
https://github.com/avehtari/BDA_R_demos