---
title: "BDA - Assignment 2"
author: "Anonymous"
date: "7/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Default eco=TRUE
```

```{r, include=FALSE} 
# Packages
library(bayesplot)
#install.packages("bayestestR")
#remotes::install_github("avehtari/BDA_course_Aalto", subdir = "rpackage")
library(aaltobda)
library(tidyverse)
```

## Q - Inference for binomial proportion

We use the dataset algae.txtw which contains observations form 274 Finnish Lakes, 
where a "0" means: NO algae in the lake, 
and "I" means: Algae in lake. 

```{r, include=FALSE}
#algae <- read_csv("C:/Users/majhe/Documents/GitHub/BDA_course_Aalto/exercises/additional_files/algae.txt")
algae <- read.csv("C:/Users/majhe/Documents/GitHub/BDA_course_Aalto/exercises/additional_files/algae.txt", header = FALSE)$V1
#usethis::use_data(algae, overwrite = TRUE)
```


### a) 
The unknown value $\pi$ according to the observations (n=observations =274, y=lakes with algaes =  44) and prior knowledge:    
$$ p(\pi \mid y)\propto \, Beta(\alpha + y, \, \beta-y+n) \, = \,Beta(2 + 44, \, 10+274-44)\, = \, Beta(46, \, 240)$$

We now want to create a model "beta_point_est" that can calculate the posterior mean

```{r, echo=TRUE}
# point estimate

beta_point_est <- function(prior_alpha, prior_beta, data){
  n <- length(data) # 274 observation
  y <- length(which(data==1)) # 44 lakes with algaes
  post_mean=(prior_alpha+y)/(prior_alpha+prior_beta+n)
  return(post_mean)
}

beta_point_est(prior_alpha = 2, prior_beta = 10, data = algae)
```
We find the posterior mean to be 16.08 %  


We now want to create a model "beta_interval" that can find the 90 % posterior interval (credible interval)

```{r, echo=TRUE}
# posterior interval using qbeta (quantile function)

beta_interval <- function(prior_alpha, prior_beta, data, prob){
  n <- length(data) 
  y <- length(which(data==1))
  pi_low <- qbeta((1-prob)/2, prior_alpha+y, prior_beta+n-y)
  pi_high <- qbeta(prob+(1-prob)/2, prior_alpha+y, prior_beta+n-y)
  post_int <- list(pi_low, pi_high)
  return(post_int)
}

beta_interval(prior_alpha = 2, prior_beta = 10, data = algae, prob = 0.9)
```
90% posterior interval: 12.66% - 19.78%

### b) 
Finding the probability that the proportion of monitoring sites with detectable algae levels $\pi$ is smaller than $\pi$(0) = 0.2 that is known from historical records by creating function "beta_low"

```{r, echo=TRUE}
# creating function beta_low with pbeta (distribution function):

beta_low <- function(prior_alpha, prior_beta, data, pi_0){
  n <- length(data) 
  y <- length(which(data==1))
  dis_fun <- pbeta(pi_0, prior_alpha+y, prior_beta+n-y)
  return(dis_fun)
}

beta_low(prior_alpha = 2, prior_beta = 10, data = algae, pi_0 = 0.2)

```
We find that the probability of that the algae level is smaller than the historical records is 95.86% 


### c) 
Assumptions for the model used in b):  

* Since we use $\beta$ functions for the priors, they must follow a $\beta$ distribution   

Add more...

### d) 
Prior sensitivity analysis:  
By testing different reasonable priors a sensitivity analysis will be carried out

```{r, echo=TRUE}
# uniform prior (1, 1):
post_mean_1.1 <- beta_point_est(prior_alpha = 1, prior_beta = 1, data = algae) # posterior mean
post_int_1.1 <- beta_interval(prior_alpha = 1, prior_beta = 1, data = algae, prob = 0.9) # 90% posterior interval
post_mean_1.1; post_int_1.1

# uniform prior (2, 2) - weak/uninformative prior:
post_mean_2.2 <- beta_point_est(prior_alpha = 2, prior_beta = 2, data = algae) # # posterior mean
post_int_2.2 <- beta_interval(prior_alpha = 2, prior_beta = 2, data = algae, prob = 0.9) # 90% posterior interval
post_mean_2.2; post_int_2.2

# uniform prior (2, 2) - weak/uninformative prior:
post_mean_2.2 <- beta_point_est(prior_alpha = 2, prior_beta = 2, data = algae) # # posterior mean
post_int_2.2 <- beta_interval(prior_alpha = 2, prior_beta = 2, data = algae, prob = 0.9) # 90% posterior interval
post_mean_2.2; post_int_2.2

# String prior (50, 150) - weak/uninformative prior:
post_mean_50.150 <- beta_point_est(prior_alpha = 50, prior_beta = 150, data = algae) # # posterior mean
post_int_50.150 <- beta_interval(prior_alpha = 50, prior_beta = 150, data = algae, prob = 0.9) # 90% posterior interval
post_mean_50.150; post_int_50.150
```

With a weak prior the data is dominant in the posteior.  
With a strong prior the posterior interval is more narrow.  


