---
title: "BDA - Assignment 6"
author: "Anonymous"
output: pdf_document
---

```{r setup, include=FALSE}
# This chunk just sets echo = TRUE as default (i.e. print all code)
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse) 
library(aaltobda)
library(rstan)
# stan settings:
source('stan_utility.R') # diagnosis of rhats
options(mc.cores = parallel::detectCores()) #for local computer
rstan_options(auto_write = TRUE) # autosave Stan
# bay settings:
library(loo) #pred. error of MCMC log likelihood
library(gridExtra)
library(bayesplot) #plots of posterior draws (mcmc_hist etc)
library(shinystan) # model paramteres & MCMC simulations
bayesplot_theme_set() #default
SEED <- 48927 # random seed for reproducability
data("bioassay")
```

## Q1

With the bioassay data from the aaltobda package, we will create a stan model, that replicate the computations in section 3.7 in BDA3.
we base it on the Gaussian prior from assignment 4 and 5, where:

$\begin{bmatrix} \alpha \\ \beta\end{bmatrix} \sim \mathcal{N} \{\mathbf{\mu_0}, \mathbf{\Sigma_0} \}$, $\mathbf{\mu_0} = \begin{pmatrix} \mu_{\alpha} \\ \mu_{\beta} \end{pmatrix} =  \begin{pmatrix} 0 \\ 10 \end{pmatrix}$ and $\mathbf{\Sigma_0} = \begin{pmatrix} \sigma_{\alpha}^2 & \rho \sigma_{\alpha} \sigma_{\beta}\\ \rho \sigma_{\alpha} \sigma_{\beta} & \sigma_{\beta}^2 \end{pmatrix} = \begin{pmatrix}  4 & 10\\ 10 & 100 \end{pmatrix}$ .

The created stanmodel: 

```{r}
# display the stanmodel "ex6.stan"
writeLines(readLines("ex6.stan"))
```

Before calling our model, we need to summarize the data into a vector:
```{r}
N = length(bioassay$x)
x = bioassay$x # dose
y = bioassay$y # deaths
n = bioassay$n # animals
# values
mu_alpha = 0
mu_beta = 10
sigma_alpha = 2
sigma_beta = 10
corr = 0.5 
# the sigma matrix:
sigma <-  matrix(c(sigma_alpha^2, corr*sigma_alpha*sigma_beta,
                   corr*sigma_alpha*sigma_beta, sigma_beta^2 ), 
                 ncol=2)
mu = c(mu_alpha, mu_beta)
# binomial data list
d_bin <- list(N = N,
              n = n,
              x = x,
              y = y, 
              sigma_theta = sigma, 
              mu_theta = mu)
```

We can now fit the data to our stan model:
```{r}
fit_bioassay <- stan(file="ex6.stan", data = d_bin, seed = SEED)
```

## Q2

For convergence analysis, we can use the build-in $\hat{R}$ analysis in rstan:

```{r}
# Monitor takes an array of simulations as it argument
# probs: specifying quantiles of interest
monitor(fit_bioassay, probs = c(0.1, 0.5, 0.9))
print(fit_bioassay) #returned by Stan's sampling function
```

Above we can find the $\hat{R}$ values to:

+ $\alpha$ (=theta[1]) = 1
+ $\beta$ (=theta[2]) = 1.

Since the $\hat{R}$ values for $\alpha$ and $\beta$ are both 1 we conclude that the model has converged. This means that our estimates are acceptable as they get closer and closer to the real value as the iterations proceeds.

## Q3

Scatter plot of $\alpha$ and $\beta$ with ggplot:

```{r}
draws <- as.data.frame(fit_bioassay)
xl <- c(-3, 5)
yl <- c(-5, 35)
ggplot(data = data.frame(draws$`theta[1]`, draws$`theta[2]`)) +
  geom_point(aes(draws$`theta[1]`, draws$`theta[2]`), color = 'darkgreen', size = 0.4) +
  coord_cartesian(xlim = xl, ylim = yl) +
  labs(x = 'alpha', y = 'beta') +
  ggtitle("Samples")
```

We can also use some of the mcmc plot functions to visualize $\alpha$ (theta[1]) and $\beta$ (theta[2]):
```{r}
# histogram of alpha and beta,
# plots marginal posterior distributions combining all chains
color_scheme_set("green")
p1 <- mcmc_hist(draws, pars = 'theta[1]', binwidth = NULL)
p2 <- mcmc_hist(draws, pars = 'theta[2]', binwidth = NULL)
grid.arrange(p1, p2, ncol=2)
# density
mcmc_dens(draws, pars = c("theta[1]", "theta[2]"),
          facet_args = list(nrow = 2))
```

**References:**

Based on code examples from:  
https://github.com/avehtari/BDA_R_demos