---
title: "BDA - Assignment 5"
author: "Anonymous"
date: "22/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Default eco=TRUE
# if echo=FALSE: Code is not included
```
```{r, include=FALSE}
library(mvtnorm)
library(aaltobda)
library(gridExtra)
library(tidyverse)
# using the bioassay dataset from aaltobda package
data("bioassay")
data <- bioassay
```

## Bioassay with Metropolis
Replicating the Metropolis algorithm for the Bioassay dataset

## Q1 - Implementing the Metropolis algorithm as an R function for the bioassay dataset

We use the following Gaussian prior: 
$$
\begin{bmatrix} \alpha \\ \beta\end{bmatrix} \sim \mathcal{N} \{\mathbf{\mu_0}, \mathbf{\Sigma_0} \}
$$
Where 
$$ \mathbf{\mu_0} = \begin{pmatrix} \mu_{\alpha} \\ \mu_{\beta} \end{pmatrix} =  \begin{pmatrix} 0 \\ 10 \end{pmatrix}$$  
and  
$$ \mathbf{\Sigma_0} = \begin{pmatrix} \sigma_{\alpha}^2 & \rho \sigma_{\alpha} \sigma_{\beta}\\ \rho \sigma_{\alpha} \sigma_{\beta} & \sigma_{\beta}^2 \end{pmatrix} = \begin{pmatrix}  4 & 10\\ 10 & 100 \end{pmatrix}  $$ .

### a) 
Implementing the density rate: 

```{r}
# From ex 4 we have our p_log_prior and p_log_posterior functions:

mu_alpha <- 0; sd_alpha <- 2
mu_beta <- 10; sd_beta <- 10 
rho <- 0.5 # correlation

p_log_prior <- function(alpha, beta){
  x = cbind(alpha, beta)
  d <- dmvnorm(x, mean=c(mu_alpha, mu_beta), 
               sigma = matrix(c(sd_alpha^2, rho*sd_alpha*sd_beta, rho*sd_alpha*sd_beta, sd_beta^2 ),
                              ncol=2))
  return(log(d))
}

p_log_posterior <- function(alpha, beta, x=bioassay$x, y=bioassay$y, n=bioassay$n){
  p_log_likelihood <- bioassaylp(alpha, beta, x, y, n)
  post <- p_log_prior(alpha, beta) + p_log_likelihood
  return(post)
}

# we will use these in our density_ratio:
density_ratio <- function(alpha_propose, alpha_previous,
                          beta_propose, beta_previous,
                          x = bioassay$x, y = bioassay$y, n = bioassay$n)
  {
  p1 <- p_log_posterior(alpha_propose, beta_propose, x=bioassay$x, y=bioassay$y, n=bioassay$n)
  p0 <- p_log_posterior(alpha_previous, beta_previous, x=bioassay$x, y=bioassay$y, n=bioassay$n)
  ratio <- exp(p1 - p0) 
  return(ratio)
}

# test, result should be 0.84
density_ratio(alpha_propose = 0.374, alpha_previous = 1.89, beta_propose = 20.04, beta_previous = 24.76, 
              x = bioassay$x, 
              y = bioassay$y, 
              n = bioassay$n)
```


### b) Implementing the metropolis algorithm using denstity ratio:

Creating a function "metropolis_bioassay()" that implements the metropolis algorithm.
We use a normal proposal distribution from the text:
$$\alpha^* \sim \mathcal{N} \{ \alpha_{t-1}, \sigma=1 \} $$ 
and 
$$\beta^* \sim \mathcal{N} \{ \beta_{t-1}, \sigma=5 \}$$. 

```{r}
# The proposal distribution:
proposal_distribution <- function(param){
  sigma = matrix(c(1, 2, 2, 5), ncol=2)
  return(rnorm(2, mean = param, sd = sigma))
}

# The Metropolis function:
metropolis_bioassay <- function(startvalue, iterations){
  chain = array(dim = c(iterations+1,2))
  chain[1,] = startvalue
  for (i in 1:iterations){
    proposal <- proposal_distribution(chain[i,])
    r <- density_ratio(alpha_propose = proposal[1], alpha_previous = chain[i,1],
                       beta_propose = proposal[2], beta_previous = chain[i,2],
                       x = bioassay$x, y = bioassay$y, n = bioassay$n)
    if (runif(1) < r){
      chain[i+1,] = proposal
    }else{
      chain[i+1,] = chain[i,]
    }
  }
  return(chain)
}
```

## Q2 - Metropolis continued..

Including the number of chains used, the starting pints, tje number of draws and the warm-up length:

```{r}
n_chains <- 5 # numbers of chain
iterations <- 50000 # number of draws
startvalues <- matrix(c(-1,1,-5,5,-2,2,4,-5,10,-1), 
                      byrow = T, ncol=2)
warm_up <- iterations/2 #warmup length

run_in_chains <- function(n_chains, iterations, startvalues){
  m <- n_chains*2
  n <- round(iterations/2/2)
  chains <- array(dim = c(2*n,2, n_chains))
  for(j in 1:n_chains){
    chain <- metropolis_bioassay(startvalues[j,], iterations)
    warm_up <- iterations/2
    new_chain <- chain[-(1:warm_up),]
    new_chain <- new_chain[1:warm_up,]
    chains[, ,j] <- new_chain
  }
  return (chains)
}
  
chains <- run_in_chains(n_chains, iterations, startvalues)

```

Missing the plot...

## Q3 - Convergence analysis

$\hat{R}$ for convergence analysis using the BDA3 version (11.4), where $\hat{R}$ compares the between- and within-chain estimates for model parameters

```{r}
R_hat <- function(chains){
  m <- 2*dim(chains)[3]
  n <- round(dim(chains)[1]/2)
  psi_ij <- array(dim = c(n, 2, m))
  for(j in 1:n_chains){    
    psi_ij[, , 2*j-1] <- chains[(1:n), , j]
    psi_ij[, , 2*j] <- chains[-(1:n), , j]
  }
  
  psi_j <- array(dim = c(m, 2))
  s_j_2 <- array(dim = c(m, 2))
  for(j in 1:m){
    psi_j[j,] <- colMeans(psi_ij[, ,j])
    s_j_2[j,] <- rowSums(apply(psi_ij[, ,j], 1, function(x) x-psi_j[j,])^2) / (n-1)
  }
  psi <- colMeans(psi_j)
  
  B <- ( n/(m-1) ) * sum(apply(psi_j, 1, function(x) x-psi)^2)
  W <- colMeans(s_j_2)
  var_hat <- ((n-1)/n) * W  + (1/n) * B 
  R_hat <- sqrt(var_hat/W)
  return(R_hat)
}
  
```


## Q4 - Plots
Scatter plotting the draws for $\alpha$ and $\beta$ 

```{r}
samp_A <- chains[, 1, 1]
samp_B <- chains[, 2, 1] 
xl <- c(-3, 8); yl <- c(-5, 35)
ggplot(data = data.frame(samp_A, samp_B)) +
  geom_point(aes(samp_A, samp_B), color = 'darkgreen', size = 0.2) +
  coord_cartesian(xlim = xl, ylim = yl) +
  labs(x = 'alpha', y = 'beta')
```