---
title: "BDA - Assignment 9"
author: "Anonymous"
date: "5/4/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Default eco=TRUE
# if echo=FALSE: Code is not included
```
```{r}
library(tidyr) 
library(rstan) 
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
library(loo)
library(ggplot2)
library(gridExtra)
library(bayesplot)
theme_set(bayesplot::theme_default(base_family = "sans"))
library(shinystan)
source('stan_utility.R')
library(aaltobda)
SEED <- 48927 # set random seed for reproducability
```




```{r}
data("factory")
```

The data applied to the hierarchical model is: 
```{r}
data_vec <-list(N = ncol(factory) * nrow(factory),
                     K = ncol(factory) + 1, # +1 is for the posterior predictive for the 7th machine
                     x = rep(1:ncol(factory), nrow(factory)),
                     y = c(t(factory)))
```

## Hierarchical model 

Mean and standard deviation:
the means of the different machines are assumed to have a common standard deviation $\sigma$ and 
The means $\mu$ are drawn from a normal distribution with $\mu_0$ and $\sigma_0$. 

Priors:
Weakly informative priors for $\mu_0$ and $\sigma_0$ and $\sigma$. 

Our hierarchical stan model:
```{r}
writeLines(readLines("hierarchical.stan"))
```

Fitting the model: 
```{r}  
fit_hierarchical <- stan(file="hierarchical.stan", data = data_vec, seed = SEED)
```

# Q1: Compute the expected utility of a product from each of the 6 machines.

Following is given:
- cost of a machine: 106 EUR
- if quality > 85, then the customers pay 200 EUR 
- if quality < 85, the prouct is not sold.
This will be applied to the utility function

Computation of the expected utility from each machine:
```{r}
Utility <- function(draws = y_pred) {
  total_profit = 0
  for(i in 1:length(draws)){
    total_profit = total_profit - 106 #minus cost of one machine
    if(draws[i] > 85) #add profit of 200 if quality >85
      total_profit = total_profit + 200 
  }
  return(total_profit/length(draws)) #divide with amount of products
}
#TEST:
y_pred <- c(83.25, 112.12, 110.20, 72.18, 82.59)
print('test of expected utility')
print(Utility(y_pred))
```

Computing the utility using the posterior predictive from each machine:
```{r}
sample_hierarchical <- rstan::extract(fit_hierarchical)
utility = c()
for(i in 1:6){
  print(sprintf('Expected utility for machine %d:',i))
  utility[i] = Utility(sample_hierarchical$ypred[,i])
  print(utility[i])
}
```


# Q2: Ranking the machines based on the expected utilities.
```{r}
order(utility)
```
according to the order the best machine is:


# Q3: A machine 7 is added: Compute and report the expected utility

The 7th machine will be based on posterior predicetion:
```{r}
utility_7 = Utility(sample_hierarchical$ypred[,7])
print('Expected utility for machine 7')
print(utility_7)
```


# Q4: Should the factory invest in a 7th machine?

No, since the expected utility is less than 0