---
title: "BDA - Assignment 4"
author: "Anonymous"
date: "8/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # Default eco=TRUE
# if echo=FALSE: Code is not included
```
```{r, include=FALSE}
library(tidyverse)
library(aaltobda)
library(mvtnorm)
library(gridExtra)
data("bioassay")
data <- bioassay
```

## Q - Bioassay model and importance sampling

### a) 
Constructing a bivariate normal distribution as a prior distribution for ($\alpha$,$\beta$):

$$
p(\alpha, \beta \mid y) \propto p(\alpha, \beta)p(\alpha, \beta \mid y)  
$$

Given is:   
$\alpha \sim \mathcal{N}(0,2^2)$  
$\beta \sim \mathcal{N}(10,10^2)$  
$corr(\alpha, \beta)=0.5$  

Specifying the values given above:    
$k=2$ and $\mu_{\alpha} =0$, $\sigma_{\alpha}=2$, $\mu_{\beta}=10$, $\sigma_{\beta}=10$ and $\rho= corr(\alpha,\beta)=0.5$. 

From this we can find $\mu$:  
$$ \mathbf{\mu} = \begin{pmatrix} \mu_{\alpha} \\ \mu_{\beta} \end{pmatrix} =  \begin{pmatrix} 0 \\ 10 \end{pmatrix}$$  
The covariance matrix is:  
$$ \mathbf{\Sigma} = \begin{pmatrix} \sigma_{\alpha}^2 & \rho \sigma_{\alpha} \sigma_{\beta}\\ \rho \sigma_{\alpha} \sigma_{\beta} & \sigma_{\beta}^2 \end{pmatrix} = \begin{pmatrix}  4 & 10\\ 10 & 100 \end{pmatrix}  $$ .


### b)
Implementing a function "p_log_prior" for computing logarithm of the density of the prior distribution of $\alpha$ $\beta$. We use the dmvnorm from the aaltobda package:

```{r}
mu_alpha <- 0
sd_alpha <- 2
mu_beta <- 10
sd_beta <- 10
rho <- 0.5 #correlation
p_log_prior <- function(alpha, beta){
  x = cbind(alpha, beta)
  d <- dmvnorm(x, mean=c(mu_alpha, mu_beta), 
               sigma = matrix(c(sd_alpha^2, rho*sd_alpha*sd_beta, rho*sd_alpha*sd_beta, sd_beta^2 ), ncol=2))
  return(log(d))
}

```


### c)
Implementing a function "p_log_posterior" for computing logarithm of the density of the posterior for arbitrary values of $\alpha$ $\beta$ and data x, y and n. Applied on dataset 'Bioassay'

The Bioassay data provided describes the outcome of toxicity tests in animals:
```{r}
head(data)
```
Where  
n = number of animals for each observation
x = dose level for each observation (log g/ml)  
y = number of deaths for each observation

```{r}
p_log_posterior <- function(alpha, beta, x=bioassay$x, y=bioassay$y, n=bioassay$n){
  p_log_likelihood <- bioassaylp(alpha, beta, x, y, n)
  post <- p_log_prior(alpha, beta) + p_log_likelihood
  return(post)
}
alpha <- 3
beta <- 9
p_log_posterior(alpha, beta, x = bioassay$x, y = bioassay$y, n = bioassay$n)
```

### d)
Plotting the posterior density in a grid of points using "bioassay_posterior_density_plot" from aaltobda package:  

```{r}
# The plot is not really working for me with the functions
# Some bug I can't find
#bioassay_posterior_density_plot(alpha_limits = c(-4, 4),
#                                beta_limits = c(-10, 30)) 
```


### e)
Two functions for 1) computing the log importance ratios.. and for 2) exponentiating the log importance ratios and normalizing them to sum to one

### function 1: 
```{r}
log_importance_weights <- function(alpha, beta){
  S <- length(alpha)
  w <- c()
  for(i in 1:S){
    w[i] <-  p_log_posterior(alpha[i], beta[i], x=bioassay$x, y=bioassay$y, n=bioassay$n) - p_log_prior(alpha[i], beta[i]) 
}  
  return(w)
} 
```

### function 2:
We divide the importance weights with the sum of importance weights, for normalization:
 
```{r}
normalized_importance_weights <- function(alpha, beta){
  log_w <- log_importance_weights(alpha, beta)
  exp_w <- exp(log_w)
  return(exp_w/sum(exp_w))
}
```

### f)
Sample draws of $\alpha$ and $\beta$ from the prior distribution

```{r}
nr <- 2000
r <- rmvnorm(nr, mean=c(mu_alpha, mu_beta),
             sigma = matrix(c(sd_alpha^2, rho*sd_alpha*sd_beta,
                              rho*sd_alpha*sd_beta, sd_beta^2 ), ncol=2))
alpha <- r[, 1]
beta <- r[, 2]
posterior_mean <- function(alpha, beta){
  ab <- cbind(alpha, beta)
  colnames(ab) <- NULL
  post <- colSums( ab * normalized_importance_weights(alpha, beta) )
  return(post)
}
```

Posterior mean of $\alpha$ and $\beta$ using importance sampling is: 

```{r}
round(posterior_mean(alpha, beta), digit=3)
```

### g)
Using importance ratio to calculate the effective sample size:

```{r}
S_eff <- function(alpha, beta){
  s_eff <- 1/sum(normalized_importance_weights(alpha, beta)^2)
  return (s_eff)
}
```

The effective sample size, S_eff, is: 
```{r}
S_eff(alpha, beta)
```

### h)
Using importance ratio to obtain a posterior sample..:

```{r}
nr <- 2000
c_R <- rmvnorm(nr, mean=c(mu_alpha, mu_beta),
             sigma = matrix(c(sd_alpha^2, rho*sd_alpha*sd_beta,
                              rho*sd_alpha*sd_beta, sd_beta^2 ), ncol=2)) 
c_A <- c_R[, 1]
c_B <- c_R[, 2]
n_s <- 1000
s_indices <- sample(length(c_A), size = n_s, replace = FALSE, 
                       prob = normalized_importance_weights(c_A, c_B) )
s_A <- c_A[s_indices[1:n_s]]
s_B <- c_B[s_indices[1:n_s]]
xl <- c(-4, 6)
yl <- c(-10, 31)
```

Scatterplot of the obtained posterior sample
```{r}
ggplot(data = data.frame(s_A, s_B)) +
  geom_point(aes(s_A, s_B), color = 'blue', size = 0.3) +
  coord_cartesian(xlim = xl, ylim = yl) +
  labs(x = 'alpha', y = 'beta')
```

### i)
Investigating if the drug is harmful by using the importance resampling without replacement:

```{r}
bpi <- s_B > 0
s_ld50 <- -s_A[bpi]/s_B[bpi]
p_positive_beta <- length(bpi)/n_s
```

Harmful drug is given by:  
$$p(\beta>0 | n,x,y)$$ 
```{r}
p_positive_beta
```


### j) 
Using the posterior sample obtained via importance resampling, draw a histogram of the draws from the posterior distribution. 

LD50 = Lethal dosage of 50% =  $-\alpha/\beta$.   
We calculate this for all of the posterior samplesand.

Histogram of LD50 is shown in the following figure. 
```{r}
ggplot() +
  geom_histogram(aes(s_ld50), binwidth = 0.05,
                 fill = 'darkgreen', color = 'black') +
  coord_cartesian(xlim = c(-0.8, 0.8)) +
  labs(x = 'LD50')
```